### 2016-10-10 Manuscript/SW_early_life_telomeres.Rmd
Edits
Home edits

###Abstract
1. Understanding the costs of individual early-life experiences is fundamental to understanding life-history evolution and population dynamics. Telomeres, the protective caps at the ends of chromosomes, shorten in response to oxidative stress, and telomere shortening is correlated with reduced survival. Thus, telomere dynamics may help us quantify individual variation in early-life costs, and enhance our understanding of how poor conditions in early life are related to later-life survival.

2. We tested how telomere dynamics are related to spatio-temporal variation in early-life conditions and later-life survival in the Seychelles warbler (*Acrocephalus sechellensis*), using a large dataset of `r nrow(juv)` juveniles from multiple cohorts spanning `r max(juv$LayYear)-min(juv$LayYear)` years.

@@

4. Juvenile telomere length varied markedly among cohorts, but we found no evidence that this variation was related to temporal variation in the environment. At the individual found weak evidence that territory quality was negatively related to telomere length, but the majority of variation in juvenile telomere length and dynamics was unexplained. We found no relationship between early-life telomere dynamics and survival to adulthood or later in life.

5. Our comprehensive, long-term data suggest that although telomeres shorten with age, in natural populations telomere dynamics vary enormously over both space and time. Furthermore, Ascertaining what explains this variation will be a difficult but worthwhile task, which will require combining long-term ecological and quantitative genetic data.

@@

>>>>

###Abstract
1. Understanding the costs individuals and populations face throughout their lifetimes is fundamental to understanding life-history evolution and population dynamics. Telomeres, the protective caps at the ends of chromosomes, shorten in response to oxidative stress, and telomere shortening is correlated with reduced survival. Thus, telomere dynamics may help us quantify individual variation in costs, and enhance our understanding individual variation in natural populations.

2. We tested how spatiotemporal in longitudinal telomere dynamics are related to ecology and survival in the Seychelles warbler (*Acrocephalus sechellensis*), using a large dataset of `r nrow(dd)` samples from multiple cohorts spanning `r max(juv$LayYear)-min(juv$LayYear)` years.

@@

4. Temporal length and rates of shortening varied markedly both within and among cohorts. Aside from age-related measurements, the only significant ecological predictor of telomer length  was temporal variation in food availability. We also found that cohort and age effects led to spurious associations between telomere dynamics and survival.

5. Our comprehensive, long-term data suggest that although telomeres shorten with age, in natural populations telomere dynamics vary enormously over both space and time, and that this variation cannot be attributed entirely to measurment error. Ascertaining what explains this variation will be a difficult but worthwhile task, and will require combining long-term ecological and quantitative genetic data.

@@

dd$LogAge <- log(dd$Agemonths)
m1 <- lmer(RTL~LogAge+(1|BirdID),data=dd)
m2 <- lmer(RTL~LogAge+(1|BirdID)+(1|TerritoryID),data=dd)
m3 <- lmer(RTL~LogAge+(1|BirdID)+(1|LayYear),data=dd)
m4 <- lmer(RTL~LogAge+(1|BirdID)+(1|CatchYear)+(1|LayYear),data=dd)

anova(m1,m2,m3,m4)
